A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-17 manuscript.Rmd
rewrite introduction to PAC closes #151 adress some points in #149

<!--AB this is a too steep intro. please add two lines on what traditional preregistration is and why we would like to do it (including references to classic papers on pregistration and its benefits. What are the problems (ambiguity)? This should be said before we offer a solution. I moved some parts from below to here-->

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019, p. 2:

> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc. This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses. [...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses. [...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al. 2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998). Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
<!--APA requires no line number, but pagenumber or section heading-->

Here we propose to use reproducible research projects with the full analysis code written in advance to data collection as comprehensive and unambiguous form of preregistration.
We term this approach _preregistration as code (PAC)_.
We argue that a PAC is an excellent planning tool that offers several benefits over traditional preregistration.
For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
In this document, the results are initially based on simulated data that mimicks the expected output.
We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, that is, when planning a study.

@@

We consider three criteria when planning a study.
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; the plan should only require the necessary amount of resources.

@@

## Why are PAC and reproducibility related? <!--AB maybe turn this into a subheading? Then, please think what other subsections under PAC deserve a subheading - I added some suggestions-->
Preregistrations are plagued by non-reproducibility like any other research outcome.
@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
We believe it is easier with PAC to detect missing steps or ambiguous decisions.
Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.

>>>>

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
For an a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
Preregistrations, however, have a problem.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates, like the ones supplied by [The Open Science Framework](https://osf.io/).
The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
But these templates can not remove researchers degrees of freedoms completely, because it is impossible to describe every minute detail of an analysis for all but the simplest analysis with natural language.

Our proposed solution is simple.
Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
PAC has the potential to all but eliminate researchers degrees of freedom.

We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.

## Why are PAC and reproducibility related?

First, preregistrations are plagued by non-reproducibility like any other research outcome.
@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis (they did not try to reproduce these in practice).
Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).

Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
This could change when combining the here proposed workflow and PAC.
We want to extend PAC from preregistering code to preregistering *the whole manuscript as dynamic document*.
The idea is as follows.
For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
In this document, the results are initially based on simulated data that mimics the expected output.

Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in how to write a scientific manuscript.
It follows that for preregistration they should stick to widely accepted standards of scientific writing [e.g. @apa7].
We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
We believe it is easier with PAC to detect missing steps or ambiguous decisions.
Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.

@@

Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
We consider three criteria when planning a study.
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; the plan should only require the necessary amount of resources.

@@
